---
title: "International Environmental Treaties"
author: "Thuy Nguyen"
date: "February 7, 2020"
output: html_document
---

Hi reviewers! Please allow me to work on this for 2 more days (Thurs and Fri). There's still too many things I need to figure out.  I'll let you know when I'm more ready. Thank you for your understanding. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rio)
library(here)
library(tidyverse)
library(ggplot2)
library(gghighlight)
library(colorblindr)
library(gt)

```

# Data

```{r load data}
# Raw data

data <- import(here("data", "iea_measonly.dta"))

# clean data
data <- data %>% 
  select("tsig_yr","teif_yr","subject_group", "region_continent", "orgauspices") %>% 
  filter(subject_group != "Uncoded")  %>% 
  filter(subject_group != "") %>% 
  arrange(tsig_yr)

# table showing data set using to produce graphs 
data %>% 
  head(n = 10) %>% 
  gt() %>% 
  cols_label(tsig_yr = "Year signed",
             teif_yr = "Year ratified",
             subject_group = "Group type of treaty",
             region_continent = "Continent", 
             orgauspices = "Auspice Organization") %>% 
  tab_header(title = "Treaty signed and ratified in the world between 1857 - 2016") %>% 
  tab_source_note(source_note = md("Data from [Ronald B. Mitchell. 2002-2020. International Environmental Agreements Database Project (Version 2020.1)](http://iea.uoregon.edu/)"))
```


```{r}

data %>% 
  count(subject_group) %>% 
  ggplot(aes(fct_reorder(subject_group), n), fill = subject_group) +
           geom_bar(position = "dodge") +
  facet_wrap(~agreement_type) +
  coord_flip() +
  guides(fill = FALSE) +
  labs(
    x = "Type of treaty group",
    y = "Number of treaties",
    title = "Type of treaties by subjects"
  )
```

# Mapping data
Type of treaty across continents. Tab showing each type of treaty at a time. 

``` {r}
data %>% 
  
  filter(region_continent == c("Africa", "Asia", "Central America", 
                              "Europe", "North America", "Oceania",
                              "South America")) %>% 
  # arrange(tsig_yr) %>% 
  # group_by(subject_group) %>% 
  ggplot(aes(subject_group, fill = subject_group)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  guides(fill = FALSE) +
  facet_wrap(~region_continent) + 
  scale_color_OkabeIto() +
  labs(
    x = "Type of treaty group",
    y = "Number of treaty",
    title = "Type of treaty signed by continent"
  )

map <- map_data("world") # need a map for continents

install.packages("rnaturalearth")
library(rnaturalearth)
countries <- ne_countries(returnclass = "sf") # join in with other file
ggplot(countries) +
  geom_sf()
```

# From signature to entry into force
Some signed treaties never got ratified, which is coded as a very large value in `tsig_yr`. We'll have diffrent visualization for group of treaties that were ratified and those were not. 

```{r}

dp3 <- data %>% 
  select("subject_group", "tsig_yr", "teif_yr") %>% 
  mutate(year_gap = teif_yr - tsig_yr) 

# subset of data that the gap is > 20 years (not ratified)

# 

dp3_pd <- dp3 %>% 
  mutate(ratified = ifelse(year_gap > 20, "Not Ratified", "Ratified")) %>% 
  count(subject_group, ratified) %>% 
  group_by(ratified) %>% 
  mutate(percent = (n/sum(n))*100) #%>% 

p1 <- dp3_pd %>% 
  filter(ratified == "Not Ratified") %>% 
ggplot(aes(fct_reorder(subject_group, percent), percent)) +
    geom_col() +
    coord_flip() +
  labs(
    x = "Type of treaty group",
    y = "Percent",
    title = "Treaties signed but not ratified"
  ) +
  facet_wrap(~ratified)

p2 <- dp3_pd %>% 
  filter(ratified == "Ratified") %>% 
ggplot(aes(fct_reorder(subject_group, percent), percent)) +
    geom_col() +
    coord_flip() +
  labs(
    x = "Type of treaty group",
    y = "Percent",
    title = "Treaties signed but not ratified",
    caption = "Data from Ronald B. Mitchell. 2002-2020. International Environmental Agreements Database Project (Version 2020.1).
Available at: http://iea.uoregon.edu/ Date accessed: 26 February 2020"
  ) +
  facet_wrap(~ratified)

library(patchwork)
p1 + p2

#subset of data that the year gap is < 20 years (ratified)
# Summarise table of average number of years gap between signature year and ratification year of each subject-group

dp_rat <- dp3 %>% 
  filter(year_gap < 20) %>% 
  group_by(subject_group) %>% 
  summarise(mean = mean(year_gap, na.rm = TRUE)) # how to calculate mean of year-gap for each group?

library(gt)
dp_rat %>% 
  gt()
```


# Organization Auspices

What kind of treaty are more likely to be under an organization auspices?
```{r, fig.height = 30, fig.width = 10}

data %>% 
  count(orgauspices) %>% 
  View()

data %>% 
  mutate(orgauspices = ifelse(orgauspices == "", NA_character_, orgauspices)) %>% 
  mutate(orgauspices = ifelse(orgauspices =="UN Economic Commission for Europe (UNECE)" |
                              orgauspices =="UNECE",
                              "UN Economic Commission for Europe", orgauspices)) %>% 
  mutate(orgauspices = ifelse(orgauspices =="FAO; Concluded under Article XIV of the FAO Constitution" | 
                              orgauspices =="FAO Depositary" |
                              orgauspices =="FAO;  and Article VIII of the International Plant Protection Convention", 
                              "FAO", orgauspices)) %>% 
  mutate(orgauspices = ifelse(orgauspices =="Permanent Commission of the South Pacific (CPPS)",
                              "Commission of the South Pacific", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="Permanent Commission of the South Pacific (CPPS)",
                              "Commission of the South Pacific", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="Forum Fisheries Agency (FFA)",
                              "Forum Fisheries Agency", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="UN Economic Commission for Europe",
                              "United Nations", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="UNEP CMS Secretariat" | 
                              orgauspices =="UNEP Caribbean Environment Programme",
                              "UNEP", orgauspices)) %>%
  drop_na(orgauspices) %>% 
  group_by(subject_group, orgauspices) %>% 
  count(orgauspices) %>% # Number of subject_group treaty under an org? the last line seems to be the total count. How to get rid of it?
  ggplot(aes(orgauspices, n)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~subject_group, ncol = 1) +
  theme_gray(base_size = 20)

```

