---
title: "International Environmental Treaties"
author: "Thuy Nguyen"
date: "February 7, 2020"
output: html_document
---

Stephanie Gluck Peer Review 

Area of Strength:

1) I found the labeling of the X & Y axis and also title of the plot to be well done as I was able to easily undrestand what each of the visualization was trying to convey based on the included information (e.g., the visuals can stand alone)

2) The use of coord_flip, fct_reorder, and facetting made the visualization easier to read and more impactful. 

3) In reviewing Thuy's project, I found it helpful that Thuy included summary data tables that let me know a bi tmore about the data she is trying to visualize. 

What I learned from reviewing Thuy's project: 

I had not heard of the package "patchwork" before until I saw Thuy's project. I can see myself using this package in the future for patching plots together. Also I typically tend to tidy or organize my data separately before plotting them to ggplot but from looking at Thuy's code, I found that there can be times when tidying the data and directly piping it to ggplot would be useful and straightforward. 

Area of Improvement: 

1) For the Plot "Type of treaty signed by continent", I wonder if having a stacked bar graph may be useful in this case? I played with your code a bit to make the stacked bar graph (see codes below). I also wonder if there was a way to annotate the subject group direclty onto the bars themseles, although that may make the bar look very busy especialy on the continents with many different subject groups. 

2) For the two patched plots with treaties. I think the X axis on the "Ratified" plot can be taken out since it is clear once you patch the 2 figure together that there is an X axis label on the "not ratified" plot. Similarly, I think the title for each plot can be omitted "Treaties signed but not ratified/ratified" and you can probably go with one large title above the 2 plots and even a subtitle to denote that the figures are separated by Not Ratified vs. Ratified. I made the comments in the codes for parts where you can possibly take out. 

3) For the Organizaiton Auspice plot, it may be helfpul to clean the data separately and then plot it into ggplot instead of having cleaning and plotting all together in one large step as the cleaning appeared more intensive for this particular plot. For the visualization, it might be possible to show a stacked bar graph that depicts for each auspice organization, how many percent of their support goes to each of the subject groups and you can also list the total number of treaties each organiztion supported either within the plot or as a footnote (e.g., i noticed that IMO has a large number of support for pollution related treaties and FAO only supports species and fish realted treaties). 


Hi reviewers! Please allow me to work on this for 2 more days (Thurs and Fri). There's still too many things I need to figure out.  I'll let you know when I'm more ready. Thank you for your understanding. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rio)
library(here)
library(tidyverse)
library(ggplot2)
library(gghighlight)
library(colorblindr)
library(rnaturalearth)
```

# Data

```{r load data}
# Raw data

data <- import(here("data", "iea_measonly.dta"))

# clean data
data <- data %>% 
  select("tsig_yr","teif_yr","subject_group", "region_continent", "orgauspices") %>% 
  filter(subject_group != "Uncoded")  %>% 
  filter(subject_group != "") %>% 
  arrange(tsig_yr)

# table showing data set using to produce graphs 
data %>% 
  head(n = 10) %>% 
  gt() %>% 
  cols_label(tsig_yr = "Year signed",
             teif_yr = "Year ratified",
             subject_group = "Group type of treaty",
             region_continent = "Continent", 
             orgauspices = "Auspice Organization") %>% 
  tab_header(title = "Treaty signed and ratified in the world between 1857 - 2016") %>% 
  tab_source_note(source_note = md("Data from [Ronald B. Mitchell. 2002-2020. International Environmental Agreements Database Project (Version 2020.1)](http://iea.uoregon.edu/)"))
```


```{r}

data %>% 
  count(subject_group) %>% 
  ggplot(aes(fct_reorder(subject_group, n), n, fill = subject_group)) + # I changed the code slightly as the previous code was not running for me. 
           geom_bar(position = "dodge", stat = "identity") +
  #facet_wrap(~agreement_type) + #I didn't see a column for agreement type
  coord_flip() +
  guides(fill = FALSE) +
  labs(
    x = "Type of treaty group",
    y = "Number of treaties",
    title = "Type of treaties by subjects"
  )
```

# Mapping data
Type of treaty across continents. Tab showing each type of treaty at a time. 

``` {r}

#if you want another method to filter 
continent <- c("Africa", "Asia", "Central America", 
                              "Europe", "North America", "Oceania",
                              "South America")
temp2 <- data %>% 
  filter(region_continent %in% continent)



temp2 <- data %>% 
  filter(region_continent == c("Africa", "Asia", "Central America", 
                              "Europe", "North America", "Oceania",
                              "South America")) %>% 
  # arrange(tsig_yr) %>% 
  # group_by(subject_group) %>% 
  ggplot(aes(subject_group, fill = subject_group)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  guides(fill = FALSE) +
  facet_wrap(~region_continent) + 
  scale_color_OkabeIto() +
  labs(
    x = "Type of treaty group",
    y = "Number of treaty",
    title = "Type of treaty signed by continent"
  )

#SG, I played around with your data/plot a bit below. I just wonder if this is a unique case where stacked bar graph is actaully acceptable since there seems to be so few treaty signed in some continents such as Oceania and North America which a stacked bar graph can really highlight. I know for other countires like Europe and Asia, the colors that stack ontop of each other may become a bit much. I also wonder if it was possible to just annonate with text within the bars what each of the treaty is about. 

data %>% 
  filter(region_continent == c("Africa", "Asia", "Central America", 
                              "Europe", "North America", "Oceania",
                              "South America")) %>% 
  # arrange(tsig_yr) %>% 
  # group_by(subject_group) %>% 
  ggplot(aes(region_continent, fill = subject_group)) +
  geom_bar() +
  coord_flip() +
  #guides(fill = FALSE) +
  #facet_wrap(~region_continent) + 
  scale_color_OkabeIto() +
  labs(
    x = "Type of treaty group",
    y = "Number of treaty",
    title = "Type of treaty signed by continent"
  ) + 
  theme(legend.position = "bottom")


countries <- ne_countries(returnclass = "sf") %>%  # join in with other file
          right_join(temp2)
ggplot(countries) +
  geom_sf()
```

# From signature to entry into force
Some signed treaties never got ratified, which is coded as a very large value in `tsig_yr`. We'll have diffrent visualization for group of treaties that were ratified and those were not. 

```{r}

dp3 <- data %>% 
  select("subject_group", "tsig_yr", "teif_yr") %>% 
  mutate(year_gap = teif_yr - tsig_yr) 

# subset of data that the gap is > 20 years (not ratified)

dp3_pd <- dp3 %>% 
  mutate(ratified = ifelse(year_gap > 20, "Not Ratified", "Ratified")) %>% 
  count(subject_group, ratified) %>% 
  group_by(ratified) %>% 
  mutate(percent = (n/sum(n))*100) #%>% 

p1 <- dp3_pd %>% 
  filter(ratified == "Not Ratified") %>% 
ggplot(aes(fct_reorder(subject_group, percent), percent)) +
    geom_col() +
    coord_flip() +
  labs(
    x = "Type of treaty group",
    y = "Percent",
    title = "Treaties signed but not ratified"
  ) +
  facet_wrap(~ratified) #I think you might not need the facet_wrap since you only filtered by "not ratified". 

p2 <- dp3_pd %>% 
  filter(ratified == "Ratified") %>% 
ggplot(aes(fct_reorder(subject_group, percent), percent)) +
    geom_col() +
    coord_flip() +
  labs(
    x = "Type of treaty group", # Maybe take out the X axis label here since it is clear once you patch the 2 figure together that there is an X axis label on the "not ratified" plot. 
    y = "Percent",
    title = "Treaties signed but ratified", #corrected from not ratified to ratified
    caption = "Data from Ronald B. Mitchell. 2002-2020. International Environmental Agreements Database Project (Version 2020.1).
Available at: http://iea.uoregon.edu/ Date accessed: 26 February 2020"
  ) +
  facet_wrap(~ratified) # might not need to facet wrap if already filtered

library(patchwork)
p1 + p2

#subset of data that the year gap is < 20 years (ratified)
# Summarise table of average number of years gap between signature year and ratification year of each subject-group

dp_rat <- dp3 %>% 
  filter(year_gap < 20) %>% 
  group_by(subject_group) %>% 
  summarise(mean = mean(year_gap, na.rm = TRUE)) # how to calculate mean of year-gap for each group?

library(gt)
dp_rat %>% 
  gt()
```


# Organization Auspices

What kind of treaty are more likely to be under an organization auspices?
```{r, fig.height = 30, fig.width = 10}

data %>% 
  count(orgauspices) %>% 
  View()

data %>% 
  mutate(orgauspices = ifelse(orgauspices == "", NA_character_, orgauspices)) %>% 
  mutate(orgauspices = ifelse(orgauspices =="UN Economic Commission for Europe (UNECE)" |
                              orgauspices =="UNECE",
                              "UN Economic Commission for Europe", orgauspices)) %>% 
  mutate(orgauspices = ifelse(orgauspices =="FAO; Concluded under Article XIV of the FAO Constitution" | 
                              orgauspices =="FAO Depositary" |
                              orgauspices =="FAO;  and Article VIII of the International Plant Protection Convention", 
                              "FAO", orgauspices)) %>% 
  mutate(orgauspices = ifelse(orgauspices =="Permanent Commission of the South Pacific (CPPS)",
                              "Commission of the South Pacific", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="Permanent Commission of the South Pacific (CPPS)",
                              "Commission of the South Pacific", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="Forum Fisheries Agency (FFA)",
                              "Forum Fisheries Agency", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="UN Economic Commission for Europe",
                              "United Nations", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="UNEP CMS Secretariat" | 
                              orgauspices =="UNEP Caribbean Environment Programme",
                              "UNEP", orgauspices)) %>%
  drop_na(orgauspices) %>% 
  group_by(subject_group, orgauspices) %>% 
  count(orgauspices) %>% 
  ggplot(aes(orgauspices, n)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~subject_group, ncol = 1) +
  theme_gray(base_size = 20)

```

