---
title: "International Environmental Treaties"
author: "Thuy Nguyen"
date: "February 7, 2020"
output: html_document
---

Hi reviewers! Please allow me to work on this for 2 more days (Thurs and Fri). There's still too many things I need to figure out.  I'll let you know when I'm more ready. Thank you for your understanding. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rio)
library(here)
library(tidyverse)
library(ggplot2)
library(gghighlight)
library(colorblindr)
library(rnaturalearth)
library(gt)

```

# Data

```{r load data}
# Raw data

data <- import(here("data", "iea_measonly.dta"))

data <- read_dta("iea_measonly.dta")
#I don't know why, but the above code told me there was no such file in my directory, so I reset the directory and ran the below line

# clean data
data <- data %>% 
  select("tsig_yr","teif_yr","subject_group", "region_continent", "orgauspices") %>% 
  filter(subject_group != "Uncoded")  %>% 
  filter(subject_group != "") %>% 
  arrange(tsig_yr)

# table showing data set using to produce graphs 
data %>% 
  head(n = 10) %>% 
  gt() %>% 
  cols_label(tsig_yr = "Year signed",
             teif_yr = "Year ratified",
             subject_group = "Group type of treaty",
             region_continent = "Continent", 
             orgauspices = "Auspice Organization") %>% 
  tab_header(title = "Treaty signed and ratified in the world between 1857 - 2016") %>% 
  tab_source_note(source_note = md("Data from [Ronald B. Mitchell. 2002-2020. International Environmental Agreements Database Project (Version 2020.1)](http://iea.uoregon.edu/)"))

#VEry cool data! Could you drop the last column since it apparently does not contain any data?
#The title is a little misleading. "In the world" would include a lot of treaties. Do you mean "at the international level"?
#Also, the title says 1857-2016 but the most recent row is for 1884. If you only display the first 10 I would take that into account in the title or plot all (which I bet was your plan anyways)
#Can you shorten the link in the caption? It's visually a little dominating right now

```


```{r}

data %>% 
  count(subject_group) %>% 
  ggplot(aes(fct_reorder(subject_group), n), fill = subject_group) +
           geom_bar(position = "dodge") +
  facet_wrap(~agreement_type) +
  coord_flip() +
  guides(fill = FALSE) +
  labs(
    x = "Type of treaty group",
    y = "Number of treaties",
    title = "Type of treaties by subjects"
  )
```

# Mapping data
Type of treaty across continents. Tab showing each type of treaty at a time. 

``` {r}
temp2 <- data %>% 
  filter(region_continent == c("Africa", "Asia", "Central America", 
                              "Europe", "North America", "Oceania",
                              "South America")) 
  arrange(tsig_yr) %>% 
  group_by(subject_group) %>% 
  ggplot(aes(subject_group, fill = subject_group)) +
  geom_bar(position = "dodge") +
  coord_flip() +
  guides(fill = FALSE) +
  facet_wrap(~region_continent) + 
  scale_color_OkabeIto() +
  labs(
    x = "Type of treaty group",
    y = "Number of treaty",
    title = "Type of treaty signed by continent"
  )

#The code for this plot does not run, yielding the following error message:
#Error: `data` must be a data frame, or other object coercible by `fortify()`, not an S3 object with class uneval
#Did you accidentally pass `aes()` to the `data` argument?
# When I uncomment the two lines it tells me it cannot find the variable tsig_yr

countries <- ne_countries(returnclass = "sf") %>%  # join in with other file
          right_join(temp2)
ggplot(countries) +
  geom_sf()
#This doesn't run either, objects ne_countries and countries do not exist

```

# From signature to entry into force
Some signed treaties never got ratified, which is coded as a very large value in `tsig_yr`. We'll have diffrent visualization for group of treaties that were ratified and those were not. 

```{r}

dp3 <- data %>% 
  select("subject_group", "tsig_yr", "teif_yr") %>% 
  mutate(year_gap = teif_yr - tsig_yr) 

# subset of data that the gap is > 20 years (not ratified)

dp3_pd <- dp3 %>% 
  mutate(ratified = ifelse(year_gap > 20, "Not Ratified", "Ratified")) %>% 
  count(subject_group, ratified) %>% 
  group_by(ratified) %>% 
  mutate(percent = (n/sum(n))*100) #%>% 

p1 <- dp3_pd %>% 
  filter(ratified == "Not Ratified") %>% 
ggplot(aes(fct_reorder(subject_group, percent), percent)) +
    geom_col() +
    coord_flip() +
  labs(
    x = "Type of treaty group",
    y = "Percent",
    title = "Treaties signed but not ratified"
  ) +
  facet_wrap(~ratified)

p2 <- dp3_pd %>% 
  filter(ratified == "Ratified") %>% 
ggplot(aes(fct_reorder(subject_group, percent), percent)) +
    geom_col() +
    coord_flip() +
  labs(
    x = "Type of treaty group",
    y = "Percent",
    title = "Treaties signed but not ratified",
    caption = "Data from Ronald B. Mitchell. 2002-2020. International Environmental Agreements Database Project (Version 2020.1).
Available at: http://iea.uoregon.edu/ Date accessed: 26 February 2020"
  ) +
  facet_wrap(~ratified)

#Very cool plot, I love your topic! 
#It's weird to see the title duplicated in the facets, can you give a title for the overall plot?
#Can you change the background and surrounding lines for the facets? White background and no lines would be more parsimonious
#Alternatively, I wonder if you could do the total number of treaties per type and then (stacked barplot) the number that is not ratified. That would allow for a more direct comparison

library(patchwork)
p1 + p2

#subset of data that the year gap is < 20 years (ratified)
# Summarise table of average number of years gap between signature year and ratification year of each subject-group

dp_rat <- dp3 %>% 
  filter(year_gap < 20) %>% 
  group_by(subject_group) %>% 
  summarise(mean = mean(year_gap, na.rm = TRUE)) # how to calculate mean of year-gap for each group?

library(gt)
dp_rat %>% 
  gt()

#Could you mean-center the variable? That way it would be easier to interpret the data and then you could also plot the difference as bar charts plus/minus 0

```


# Organization Auspices

What kind of treaty are more likely to be under an organization auspices?
```{r, fig.height = 30, fig.width = 10}

data %>% 
  count(orgauspices) %>% 
  View()

data %>% 
  mutate(orgauspices = ifelse(orgauspices == "", NA_character_, orgauspices)) %>% 
  mutate(orgauspices = ifelse(orgauspices =="UN Economic Commission for Europe (UNECE)" |
                              orgauspices =="UNECE",
                              "UN Economic Commission for Europe", orgauspices)) %>% 
  mutate(orgauspices = ifelse(orgauspices =="FAO; Concluded under Article XIV of the FAO Constitution" | 
                              orgauspices =="FAO Depositary" |
                              orgauspices =="FAO;  and Article VIII of the International Plant Protection Convention", 
                              "FAO", orgauspices)) %>% 
  mutate(orgauspices = ifelse(orgauspices =="Permanent Commission of the South Pacific (CPPS)",
                              "Commission of the South Pacific", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="Permanent Commission of the South Pacific (CPPS)",
                              "Commission of the South Pacific", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="Forum Fisheries Agency (FFA)",
                              "Forum Fisheries Agency", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="UN Economic Commission for Europe",
                              "United Nations", orgauspices)) %>%
  mutate(orgauspices = ifelse(orgauspices =="UNEP CMS Secretariat" | 
                              orgauspices =="UNEP Caribbean Environment Programme",
                              "UNEP", orgauspices)) %>%
  drop_na(orgauspices) %>% 
  group_by(subject_group, orgauspices) %>% 
  count(orgauspices) %>% 
  ggplot(aes(orgauspices, n)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~subject_group, ncol = 1) +
  theme_gray(base_size = 20)

#This plot is not legible for me because it shows too many rows on the left and too many facets
#You could just pick a few types (like the three types with the ost treaties) and the plot would be a lot easier to understand

```

